### 2020-10-19 inst/gitbook/02_detailed_functions.Rmd
Create 02_detailed_functions.Rmd

---
title: "Getting start"
author: "Thomas Guillerme (guillert@tcd.ie)"
date: "`r Sys.Date()`"
bibliography: [../References.bib, ../packages.bib]
output:
  html_document: default
  pdf_document: default
---

# Making trait processes with `make.traits()`

Trait processes always need the following input arguments

 * `x0`: the previous trait value(s)
 * `edge.length`: the branch length value
 * `...`: a placeholder for any extra arguments

If it contains at least these arguments, it should be handled correctly by `make.traits` an `dads`

> It is highly recommended to set all arguments (but `x0`) to a default value.