### 2020-10-23 inst/gitbook/index.Rmd
Update index.Rmd

--- 
title: "dads manual"
author: "Thomas Guillerme (guillert@tcd.ie), Alex Slavenko (email) and others"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [../references.bib]
link-citations: yes
github-repo: TGuillerme/dads
description: "dads R package vignette"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# `dads`

Disparity and diversity simulations.

Allowing to simulate disparity and diversity at the same time with interaction between both.

The core of this package is based on the [`diversitree` birth-death algorithm](https://github.com/cran/diversitree/blob/master/R/simulate-bd.R).

## What is `dads`?

There are some very good packages out there to simulate birth death trees (like [TreeSim](https://github.com/tanja819/TreeSim)) or even packages simulating traits (disparity) and diversity jointly (like [RPANDA](https://github.com/cran/RPANDA) or [PETER](https://github.com/PuttickMacroevolution/PETER)).
We strongly advice you have a look at these packages first as they might be more appropriate for your tasks.

`dads` aims to be a highly modular and friendly version of all these packages: it allows to simulate disparity and diversity jointly with a vast array of options that can be easily modified by users.
For example, you can easily generate any type of process to generate a trait (BM, OU, something else, etc...) in multiple dependent or independent dimensions through `"traits"` objects.
You can then specify how the traits should affect disparity through `"modifiers"` objects. And finally you can create events (like mass extinctions) through `"events"` objects.
These objects and how to modify them will be detailed throughout this manual.
Finally we are putting an emphasise in the development of this package on the speed and reliability of the functions.

### Modular?

Because their is an infinite way you might want to generate disparity and diversity (different traits, different modifiers and different events), the `dads` package is designed to make all these parts easy to code separately and integrate them easily in the `dads` core functions.
This allows you to simulate finely tuned multidimensional data for your specific project!

## Installing and running the package

You can install this package easily, directly from the github:
<!-- TG: I do intend to release it on CRAN later on -->

```{r, eval = FALSE}
## Checking if devtools is already installed
if(!require(devtools)) install.packages("devtools")

## Installing the latest version directly from GitHub
install_github("TGuillerme/dads")
```

```{r, eval = FALSE, echo = FALSE, message = FALSE}
## loading the package and setting up the start seed.
library(dads)
## Setting a random seed for repeatability
set.seed(123)
```

## Help

If you need help with the package, hopefully the following manual will be useful.
However, parts of this package are still in development and some other parts are probably not covered.
Thus if you have suggestions or comments on on what has already been developed or will be developed, please send me an email (<guillert@tcd.ie>) or if you are a GitHub user, directly create an issue on the [GitHub page](https://github.com/TGuillerme/dads).
Doing so will not only hopefully help you but also other users since it will help improve this manual!