### 2020-12-01 inst/gitbook/03_make.modifiers.Rmd
Wrapped up branch.length example

---
title: "`make.modifiers`"
author: "Thomas Guillerme (guillert@tcd.ie)"
date: "`r Sys.Date()`"
bibliography: [../References.bib, ../packages.bib]
output:
  html_document: default
  pdf_document: default
---

# Making modifiers with `make.modifiers()` {#makemodifiers}

```{r, echo = FALSE}
set.seed(123)
```

`"modifiers"` have a similar structure than `"traits"` where you can design an object with increasing complexity, starting with the simplest modifiers that doesn't modify anything (using the default arguments):

```{r}
## Making a default modifier (no modification)
my_default_modifiers <- make.modifiers()
my_default_modifiers
```

Similarly to `"traits"` objects, `"modifiers"` are also printed by default using `print.dads`.
You can see details about what's actually in the object using `print.dads(my_default_modifiers, all = TRUE)`.
However, contrary to `"traits"`, you cannot plot `"modifiers"`.

## The branch length function (`branch.length`)

The first argument in `"modifiers"` is the branch length function (`branch.length`) this is the function that will be executed in `dads` to generate branch length.
Note that in the `dads` algorithm, branch length is not generated _directly_ but actually results of the waiting time.
In other words, the `branch.length` function just affects waiting time for all taxa present at any time in the simulation.
These taxa can then either go extinct (stopping the "growth" of it's branch length) or survive (continuing the "growth").

By default, branch length (or waiting or growth) is a randomly drawn number from an exponential distribution with the rate of the number of taxa multiplied by the speciation and extinction rate: $n \times (\lambda + \mu)$ (where $n$ is the number of taxa currently present in the simulation, $\lambda$ and $\mu$ are respectively the speciation and extinction rates).
This default function is simply called `branch.length` in `dads` and can be used as a modifier as follows:

```{r}
## Specifying the default modifier
default_modifiers <- make.modifiers(branch.length = branch.length)

## Setting some parameters for generating trees
bd_params  <- list(extinction = 0)
stope_rule <- list(max.living = 20)

## Generating a tree with the default branch length parameter
set.seed(0)
default_tree <- dads(bd.params = bd_params,
                     stop.rule = stope_rule,
                     modifiers = default_modifiers)
```

Of course, the point of the modularity here is that you can provide your own function for generating branch length.
For example, we might be interested in what our tree would look like if we'd use a simple constant branch length generation (instead of randomly drawing it from an exponential distribution).
We can do so by declaring our own `branch.length` function and adding it to a `"modifiers"` object.

```{r}
## A constant branch length generator
## (note that the output must be numeric, not integer)
constant.brlen <- function() {
    return(as.numeric(1))
}

## Creating the modifiers object
constant_modifier <- make.modifiers(branch.length = constant.brlen)

## Generating a new tree with this modifier
set.seed(0)
modified_tree <- dads(bd.params = bd_params,
                      stop.rule = stope_rule,
                      modifiers = constant_modifier)
```

And we can visualise the difference between both resulting trees:

```{r}
par(mfrow = c(1,2))
plot(default_tree,  main = "Default modifier")
plot(modified_tree, main = "Constant branch length\nmodifier")
```

It is of course to use more complex branch length modifiers that intakes different conditions and specific modification rather than simply always output a value of one.

## The condition and modify functions (`condition` and `modify`)

In the examples above, we have seen how to specify modifications to the birth death process (via `branch.length`, `selection` and `speciation`), however, one might note that these modifications are not dynamic.
In other words, throughout the process, the modifications remain constant (even if they are conditional).
It is however possible to code the modifiers

In the following section we will see how to program these functions in a `"modifiers"` before learning about other aspects of the birth death process that can be modified (the `selection` and `speciation` steps).

As shown in the example with the `branch.length` modifier, it is possible to modify a specific part of the birth death process (e.g. the change in time) but in the previous example, the `constant.brlen` modifier only included a constant modification of the process whereas it is probably more valuable to have a variable change in branch length (e.g. like the default `branch.length` modifier that is dependent on the number of taxa at the time of the simulation).
It is possible add conditions and modifications to the `"modifiers"` object through the `condition` and `modify` function.
As stated in their names, the `condition` function is indicating in which conditions to do a special action while the `modify` function is indicating what to modify (typically when a condition is met).

`condition` and `modify` are hard coded in the `branch.length` function that they concern, i.e. they are variables (functions) within the function.

```{r, eval = FALSE}
constant.brlen <- function(bd.params      = NULL,
                           n.taxa         = NULL,
                           parent.lineage = NULL,
                           trait.values   = NULL,
                           modify.fun     = NULL) {
    return(1)
}

constant_brlen <- dads(bd.params = list(extinction = 0.25),
                     stop.rule = list(max.living = 20),
                     traits    = make.traits(),
                     modifiers = make.modifiers(
                        branch.length = constant.brlen))

plot(constant_brlen, main = "constant branch length")
```

## The speciation function (`speciation`)

```{r, eval = FALSE}
random.brlen <- function(bd.params = NULL,
                         n.taxa = NULL,
                         parent.lineage = NULL,
                         trait.values = NULL,
                         modify.fun = NULL) {
    return(runif(1))
}

random.spec  <- function(bd.params = NULL,
                         n.taxa = NULL,
                         parent.lineage = NULL,
                         trait.values = NULL,
                         modify.fun = NULL) {
    return(sample(c(TRUE, FALSE), 1))
}

set.seed(1)
yule_tree <- dads(bd.params = list(extinction = 0),
                     stop.rule = list(max.living = 20),
                     traits    = make.traits(),
                     modifiers = make.modifiers(
                        branch.length = random.brlen,
                        speciation    = random.spec))

plot(yule_tree, main = "a (failing) yule(ish) tree")
```

## The condition function (`condition`)

## The modify function (`modify`)

## Combining and editing modifiers (`add`)

## Testing modifiers (`test`)

## Demo runnable

```{r, fig.height = 16, fig.width = 8}
bd.params <- list(speciation = 1, extinction = 1/3)
traits <- make.traits()
stop.rule <- list(max.taxa = 20, max.living = Inf, max.time = Inf)
modifiers <- NULL
events <- NULL
null.error <- NULL

## modifiers
condition <- function(trait.values, parent.lineage) return(parent.traits(trait.values, parent.lineage) < 0)
modify <- function(x) return(x * 20)

## Setting up the different modifiers
modify_speciation <- make.modifiers(speciation    = speciation.trait,
                                    condition     = condition,
                                    modify        = modify)
modify_brlen <- make.modifiers(branch.length = branch.length.trait,
                               condition     = condition,
                               modify        = modify)
modify_speciation_brlen <- make.modifiers(branch.length = branch.length.trait,
                                          speciation    = speciation.trait,
                                          condition     = condition,
                                          modify        = modify)

## Test normal (no modifiers)
set.seed(1)
test <- dads(bd.params, stop.rule, traits, null.error = 20)
par(mfrow = c(4,2))
plot(test$tree)
plot.dads(test, main = "random tree + trait")

## Test with modifiers
set.seed(1)
trait_table <- NULL
test <- dads(bd.params, stop.rule, traits,
                                modifiers = modify_speciation,
                                null.error = 20)
plot(test$tree)
plot.dads(test, main = "Skewed speciation")

set.seed(1)
trait_table <- NULL
test <- dads(bd.params, stop.rule, traits,
                                modifiers = modify_brlen,
                                null.error = 20)
plot(test$tree)
plot.dads(test, main = "Skewed branch length")

set.seed(1)
trait_table <- NULL
test <- dads(bd.params, stop.rule, traits,
                                modifiers = modify_speciation_brlen,
                                null.error = 20)
plot(test$tree)
plot.dads(test, main = "Skewed branch length and speciation")

```