### 2020-12-02 inst/gitbook/03_make.modifiers.Rmd
Update 03_make.modifiers.Rmd

---
title: "`make.modifiers`"
author: "Thomas Guillerme (guillert@tcd.ie)"
date: "`r Sys.Date()`"
bibliography: [../References.bib, ../packages.bib]
output:
  html_document: default
  pdf_document: default
---

# Making modifiers with `make.modifiers()` {#makemodifiers}

```{r, echo = FALSE}
set.seed(123)
```

`"modifiers"` have a similar structure than `"traits"` where you can design an object with increasing complexity, starting with the simplest modifiers that doesn't modify anything (using the default arguments):

```{r}
## Making a default modifier (no modification)
my_default_modifiers <- make.modifiers()
my_default_modifiers
```

Similarly to `"traits"` objects, `"modifiers"` are also printed by default using `print.dads`.
You can see details about what's actually in the object using `print.dads(my_default_modifiers, all = TRUE)`.
However, contrary to `"traits"`, you cannot plot `"modifiers"`.

## The branch length function (`branch.length`)

The first argument in `"modifiers"` is the branch length function (`branch.length`) this is the function that will be executed in `dads` to generate branch length.
Note that in the `dads` algorithm, branch length is not generated _directly_ but actually results of the waiting time.
In other words, the `branch.length` function just affects waiting time for all taxa present at any time in the simulation.
These taxa can then either go extinct (stopping the "growth" of it's branch length) or survive (continuing the "growth").

By default, branch length (or waiting or growth) is a randomly drawn number from an exponential distribution with the rate of the number of taxa multiplied by the speciation and extinction rate: $n \times (\lambda + \mu)$ (where $n$ is the number of taxa currently present in the simulation, $\lambda$ and $\mu$ are respectively the speciation and extinction rates).
This default function is simply called `branch.length` in `dads` and can be used as a modifier as follows:

```{r}
## Specifying the default modifier
default_modifiers <- make.modifiers(branch.length = branch.length)

## Setting some parameters for generating trees
bd_params  <- list(extinction = 0)
stope_rule <- list(max.living = 20)

## Generating a tree with the default branch length parameter
set.seed(0)
default_tree <- dads(bd.params = bd_params,
                     stop.rule = stope_rule,
                     modifiers = default_modifiers)
```

Of course, the point of the modularity here is that you can provide your own function for generating branch length.
For example, we might be interested in what our tree would look like if we'd use a simple constant branch length generation (instead of randomly drawing it from an exponential distribution).
We can do so by declaring our own `branch.length` function and adding it to a `"modifiers"` object.

```{r}
## A constant branch length generator
## (note that the output must be numeric, not integer)
constant.brlen <- function() {
    return(as.numeric(1))
}

## Creating the modifiers object
constant_modifier <- make.modifiers(branch.length = constant.brlen)

## Generating a new tree with this modifier
set.seed(0)
modified_tree <- dads(bd.params = bd_params,
                      stop.rule = stope_rule,
                      modifiers = constant_modifier)
```

And we can visualise the difference between both resulting trees:

```{r}
par(mfrow = c(1,2))
plot(default_tree,  main = "Default modifier")
plot(modified_tree, main = "Constant branch length\nmodifier")
```

It is of course to use more complex branch length modifiers that intakes different conditions and specific modification rather than simply always output a value of one.

### The allowed arguments {#allowarguments}

It is of course possible to design some more advanced function to interact with the birth death process.
You can create a function for `branch.length`, `selection` and `speciation` that involve any of the following arguments:

  * `bd.params`: a named list containing `"numeric"` values that contains the birth death parameters (at least `"speciation"` and `"extinction"`);
  * `lineage`: a named list containing the lineage data (see below).
  * `trait.values`: a `"matrix"` containing `"numeric"` values with the trait names as column names and the lineages ID as row numbers (you can use it with the function `parent.traits` to access the trait of the previous node for example).
  * `modify.fun`: a `"list"` of named `"function"` (usually passed through `condition` and `modify`).

The `lineage` list contains the following elements (missing elements are allowed):

 * `lineage$parents`: an `"integer"` vector: the list of parent lineages;
 * `lineage$livings`: an `"integer"` vector: the list of lineages still not extinct;
 * `lineage$drawn`: a single `"integer"`: the ID of the selected lineage;
 * `lineage$current`: a single `"integer"`: the selected lineage (is equal to `lineage$livings[lineage$drawn]`);
 * `lineage$n`: a single `"integer"`: the current number of non extinct lineage (is equal to `length(lineage$livings))`;
 * `lineage$split`: a `"logical"` vector: the list of splits for each lineage (`TRUE`), the number of total tips is equal to `sum(!lineage$split)`.

In general, unless you know what you're doing, you can ignore most arguments for specific modifiers since they are handled automatically within the `dads` function.
Therefore any argument can be left undeclared or missing and is always handled internally. For example, if you did not declare `n.taxa` as a function argument but are using `n.taxa` in the function, `n.taxa` will be detected and treated as a current argument automatically as set accordingly within the birth death process (e.g. `n.taxa` will be set to the current number of taxa every iteration of the process).

For example, we can create a function that increases branch length proportional to the number of species "alive" at each time of the simulation in a discrete way. I.e. for discrete numbers of taxa, the branch length increases by jumps (ten fold):

```{r}
## A more complex binned.branch.length function
increasing.brlen <- function(bd.params, lineage) {

    ## Setting the cumulated birth and death
    birth_death <- bd.params$speciation + bd.params$extinction

    ## Returning branch lengths depending on different number of taxa
    if(lineage$n <= 5) {
        return(1    * rexp(1, sum(5 * birth_death)))
    }
    if(lineage$n <= 10) {
        return(10   * rexp(1, sum(10 * birth_death)))   
    }
    if(lineage$n <= 15) {
        return(100  * rexp(1, sum(15 * birth_death)))      
    }
    if(lineage$n <= 20) {
        return(1000 * rexp(1, sum(20 * birth_death)))
    } else {
        return(1000 * rexp(1, sum(lineage$n * birth_death)))
    }
}
```

We can then create it as a `"modifiers"` object and run a new simulation:

```{r}
## Creating a modifiers
increasing_modifier <- make.modifiers(branch.length = increasing.brlen)

## Generating a new tree with this modifier
set.seed(0)
increasing_tree <- dads(bd.params = bd_params,
                        stop.rule = stope_rule,
                        modifiers = increasing_modifier)
```

And we can visualise the difference between the resulting trees:

```{r}
par(mfrow = c(1,3))
plot(default_tree,    main = "Default modifier")
plot(modified_tree,   main = "Constant branch length\nmodifier")
plot(increasing_tree, main = "Increasing branch length\nmodifier (binned)")
```

## The selection function (`selection`)

The `selection` function is used in the birth death process to know which lineage to select when running a speciation (or extinction!) event.
By default, this function randomly selects one taxon that is currently not extinct (using: `sample(1:n.taxa, 1))`.
Similarly as `branch.length` it is possible to modify this part of the birth death process.
For example, we could simply select always the last created lineage (to create a "ladder" or most asymmetric tree):

```{r}
## Our function to always select the last taxon
## (making sure it returns an integer)
select.last <- function(lineage) {
    return(as.integer(lineage$n))
}
```

> Note that here the function can only intake the [allowed arguments as described above](#allowarguments) (here `n.taxa`: the number of current living taxa).

We can then create a `"modifiers"` object the same way as before using this time the `selection` argument:

```{r}
## A modifier for selection
ladderised_modifier <- make.modifiers(selection = select.last)
## Generating a new tree with this modifier
set.seed(0)
ladderised_tree <- dads(bd.params = bd_params,
                        stop.rule = stope_rule,
                        modifiers = ladderised_modifier)
## Displaying the results
par(mfrow = c(1,2))
plot(default_tree,    main = "Default modifier")
plot(ladderised_tree, main = "Ladderising modifier")
```

Again, it is of course possible to make the modifier more complex and in combination with other elements of the tree.
For example, we can create a `"dads"` object that also creates a trait a add to it a `selection` modifier that only selects for tips with positive trait values.

```{r}
## Our function that only select taxa with positive trait values
select.positive <- function(trait.values, lineage) {
    ## Selecting the taxa with positive traits only
    positives <- as.integer(rownames(trait.values)[which(trait.values[, 1] >= 0)])

    ## Select the current taxa that descend from a node with a positive value
    positive_living <- cbind(lineage$parents, seq_along(lineage$split)
                             )[which(lineage$parents %in% positives), 2]

    ## Select one tip randomly in the ones with descendants with positive values
    return(sample(which(lineage$livings %in% positive_living), 1))
}

## Creating the modifier
positive_skew <- make.modifiers(selection = select.positive)

## Creating a (default) trait object
BM_trait <- make.traits()

## Simulate a tree and trait with no modifier
set.seed(1)
default_dads <- dads(bd.params = bd_params,
                     stop.rule = stope_rule,
                     traits    = BM_trait)

## Simulate a tree and trait with the modifier
set.seed(1)
skewed_trait_dads <- dads(bd.params = bd_params,
                          stop.rule = stope_rule,
                          traits    = BM_trait,
                          modifiers = positive_skew)

## Plotting the differences in trees and traits
par(mfrow = c(1, 2))
plot(default_dads, main = "Default trait and tree")
plot(skewed_trait_dads, main = "Skewed trait and tree")
```

## The speciation function (`speciation`)

The third function that can be used to modify the birth death process is the `speciation` function.
This one is used during the birth death process to decide whether a lineages speciates (creating a node and two new lineages) or goes extinct (creating a tip).

> Note that the `speciation` function only affects tips or nodes before the simulation reaches the `stop.rule`.
The then surviving lineages are all automatically transformed into tips.

By default, the `speciation` function is trigger a speciation even if a number randomly drawn from a uniform distribution is lower than the ratio between the speciation and the speciation and extinction parameter. If the randomly drawn number is higher, the lineage goes extinct.

```{r, eval = FALSE}
## The speciation in pseudo-code:
runif(1) < speciation/ (speciation + extinction)
```

Creating a `"modifiers"` with a `speciation` function works the same way as for `branch.length` and `selection` but the function that will be used needs to output a logical value (see table [below](#summarymodifiers)).
Once the function is created simply input your function for speciation in the modifier and run the `dads` function with that modifier:

```{r, eval = FALSE}
## Speciating or going extinct randomly
## (regardless of the extinction parameter)
random.extinct  <- function() {
    return(sample(c(TRUE, FALSE), 1))
}

## Creating the modifiers object
random_extinction <- make.modifiers(speciation = random.extinct)

## Generating a new tree with this modifier
set.seed(0)
modified_tree <- dads(bd.params = bd_params,
                      stop.rule = stope_rule,
                      modifiers = random_extinction)

par(mfrow = c(1,2))
plot(default_tree,  main = "Default modifier")
plot(modified_tree, main = "Random extinction\nmodifier")
```

Note how every lineage end up going extinct!
And again, we can make some more advanced modifiers: for example, one where a tip always goes extinct if their ancestor has a negative trait value (here we will also introduce the utility function `parent.trait` that automatically selects the trait values of the parent of the current lineage.

```{r}
## A modifier for removing tips with negative values
bias.trait <- function(trait.values, lineage) {
    if(parent.traits(trait.values, lineage) < 0) {
        ## Go extinct!
        return(FALSE)
    } else {
        ## Speciate
        return(TRUE)
    }
}

## Creating the modifier
biased_trait <- make.modifiers(speciation = bias.trait)

## Simulate a tree and trait with the modifier
set.seed(1)
biased_trait_dads <- dads(bd.params = bd_params,
                          stop.rule = stope_rule,
                          traits    = BM_trait,
                          modifiers = biased_trait)

## Plotting the differences in trees and traits
par(mfrow = c(1, 2))
plot(default_dads, main = "Default trait and tree")
plot(biased_trait_dads, main = "Biased trait and tree")
```

## Summary of the inputs and outputs for the `branch.length`, `selection` and `speciation` modifiers {#summarymodifiers}

modifier name   | accepted input (arguments) | required output (class) |
----------------|----------------------------|-------------------------|
`branch.length` | `bd.params`, `lineage`, `trait.values` | `"numeric"` |
`selection`     | `bd.params`, `lineage`, `trait.values` | `"integer"` |
`speciation`    | `bd.params`, `lineage`, `trait.values` | `"logical"` |

## The condition and modify functions (`condition` and `modify`)

In the examples above, we have seen how to specify modifications to the birth death process (via `branch.length`, `selection` and `speciation`), however, one might note that these modifications are not dynamic.
In other words, throughout the process, the modifications remain constant (even if they are conditional).
It is however possible to code the `"modifiers"` so that they can be affected by `"events"` objects (see next chapter).

To do so, you can formally declare conditions (`condition`) and modifications (`modify`) as internal functions that can then be modified my an `"events"` object.
`condition` and `modify` are hard coded in the `branch.length` function that they concern, i.e. they are variables (functions) within the function.

## Combining and editing modifiers (`add`)

## Testing modifiers (`test`)

## Demo runnable

```{r, fig.height = 16, fig.width = 8}
bd.params <- list(speciation = 1, extinction = 1/3)
traits <- make.traits()
stop.rule <- list(max.taxa = 20, max.living = Inf, max.time = Inf)
modifiers <- NULL
events <- NULL
null.error <- NULL

## modifiers
condition <- function(trait.values, parent.lineage) return(parent.traits(trait.values, parent.lineage) < 0)
modify <- function(x) return(x * 20)

## Setting up the different modifiers
modify_speciation <- make.modifiers(speciation    = speciation.trait,
                                    condition     = condition,
                                    modify        = modify)
modify_brlen <- make.modifiers(branch.length = branch.length.trait,
                               condition     = condition,
                               modify        = modify)
modify_speciation_brlen <- make.modifiers(branch.length = branch.length.trait,
                                          speciation    = speciation.trait,
                                          condition     = condition,
                                          modify        = modify)

## Test normal (no modifiers)
set.seed(1)
test <- dads(bd.params, stop.rule, traits, null.error = 20)
par(mfrow = c(4,2))
plot(test$tree)
plot.dads(test, main = "random tree + trait")

## Test with modifiers
set.seed(1)
trait_table <- NULL
test <- dads(bd.params, stop.rule, traits,
                                modifiers = modify_speciation,
                                null.error = 20)
plot(test$tree)
plot.dads(test, main = "Skewed speciation")

set.seed(1)
trait_table <- NULL
test <- dads(bd.params, stop.rule, traits,
                                modifiers = modify_brlen,
                                null.error = 20)
plot(test$tree)
plot.dads(test, main = "Skewed branch length")

set.seed(1)
trait_table <- NULL
test <- dads(bd.params, stop.rule, traits,
                                modifiers = modify_speciation_brlen,
                                null.error = 20)
plot(test$tree)
plot.dads(test, main = "Skewed branch length and speciation")

```