### 2021-01-12 inst/gitbook/04_make.events.Rmd
events and thorough example INIT (working-ish examples)

---
title: "`make.events`"
author: "Thomas Guillerme (guillert@tcd.ie)"
date: "`r Sys.Date()`"
bibliography: [../References.bib, ../packages.bib]
output:
  html_document: default
  pdf_document: default
---

# Making events with `make.events()` {#makeevents}

```{r, echo = FALSE}
set.seed(123)
```

## Conditions

## Modifications

## Some events examples

### Random mass extinction after some time

```{r}
## Simulation parameters
stop.rule <- list(max.time = 5)
bd.params <- list(extinction = 0, speciation = 1)

## 80% mass extinction at time 4
mass_extinction <- make.events(
                      target = "taxa",
                      condition = time.condition(4),
                      modification = random.extinction(0.8))

## Running the simulations
set.seed(1)
results <- dads(bd.params = bd.params,
                stop.rule = stop.rule,
                events    = mass_extinction)
## Plotting the results
plot(results, show.tip.label = FALSE)
axisPhylo()
```

### Extinction based on a trait value after some time

```{r}
## Simulation parameters
stop.rule <- list(max.time = 5)
bd.params <- list(extinction = 0, speciation = 1)

## Extinction of any tips with trait < 1 at time 4
trait_extinction <- make.events(
                      target = "taxa",
                      condition = time.condition(4),
                      modification = trait.extinction(1))

## Running the simulations
set.seed(7)
results <- dads(bd.params = bd.params,
                stop.rule = stop.rule,
                traits    = make.traits(),
                events    = trait_extinction)
## Plotting the results
plot(results)
```

### Adding a background extinction after reaching a number of living taxa

```{r}
## Simulation parameters
stop.rule <- list(max.living = 50)
bd.params <- list(extinction = 0, speciation = 1)

## Adding an extinction parameter after 30 taxa
background_extinction <- make.events(
                      condition    = taxa.condition(30),
                      target       = "bd.params",
                      modification = update.bd.params(1/3, "extinction"))

## Running the simulations
set.seed(2)
results <- dads(bd.params = bd.params,
                stop.rule = stop.rule,
                events    = background_extinction)
## Plotting the results
plot(results)
```

### Reducing speciation after some time

```{r}
## Simulation parameters
stop.rule <- list(max.time = 4)
bd.params <- list(extinction = 0, speciation = 1)

## Reducing speciation after reaching time 2
reduced_speciation <- make.events(
                      condition    = time.condition(2),
                      target       = "bd.params",
                      modification = update.bd.params(1/3, "speciation"))

set.seed(42)
no_event <- dads(bd.params = bd.params,
                 stop.rule = stop.rule)
set.seed(42)
reduced_speciation_event <- dads(bd.params = bd.params,
                                 stop.rule = stop.rule,
                                 events    = reduced_speciation)

## Plot both trees
par(mfrow = c(1, 2))
plot(no_event, main = "No event")
axisPhylo()
plot(reduced_speciation_event, main = "Reduced speciation after time 2")
axisPhylo()
```

### Changing trait process after some time

```{r}
## Simulation parameters
bd.params <- list(extinction = 0, speciation = 1)
stop.rule <- list(max.time = 6)
traits    <- make.traits()

## Create an event to change the trait process
change_process <- make.events(
                  condition    = time.condition(5),
                  target       = "traits",
                  modification = update.traits(process = OU.process))

## Run the simulations
set.seed(1)
no_change <- dads(bd.params = bd.params,
                  stop.rule = stop.rule,
                  traits    = traits)
set.seed(1)
process_change <- dads(bd.params = bd.params,
                       stop.rule = stop.rule,
                       traits    = traits,
                       events    = change_process)
## Plot the results
par(mfrow = c(1,2))
plot(no_change, ylim = c(-7, 7))
plot(process_change, ylim = c(-7, 7))
```

### Changing trait correlation after reaching a trait value

```{r}
## Set the parameters
stop.rule <- list(max.taxa = 100)
bd.params <- list(extinction = 0, speciation = 1)

## A correlated trait
traits <- make.traits(n = 2, process.args = list(Sigma = matrix(1, 2, 2)))

## Event changing a trait correlation
change_correlation <- make.events(
    condition    = trait.condition(3, absolute = TRUE),
    target       = "traits",
    modification = update.traits(process.args = list(Sigma = matrix(c(10,3,3,2),2,2))))

## Run the simulations
set.seed(8)
no_event <- dads(bd.params = bd.params,
                 stop.rule = stop.rule,
                 traits    = traits)
set.seed(8)
change_correlation <- dads(bd.params = bd.params,
                           stop.rule = stop.rule,
                           traits    = traits,
                           events    = change_correlation)
## Visual testing
par(mfrow = c(1,2))
plot(no_event, trait = 2)
plot(change_correlation, trait = 2)
# plot(no_event, trait = c(1:2))
# plot(change_correlation, trait = c(1:2), use.3D = TRUE)
```

### Event for changing a modifier

```{r}
## Set the parameters
stop.rule <- list(max.time = 4)
bd.params <- list(extinction = 0, speciation = 1)

## Using a default trait and modifier object
modifiers <- make.modifiers()
traits <- make.traits()

## New condition and new modifier (increasing speciation if trait is negative)
new.condition <- function(trait.values, lineage) {
    return(parent.traits(trait.values, lineage) < 0)
}
new.modify    <- function(x, trait.values, lineage) {
  return(x + 1)
}

## Update the modifier
change_speciation <- make.events(
    condition    = time.condition(3),
    target       = "modifiers",
    modification = update.modifiers(speciation = speciation,
                                    condition  = new.condition,
                                    modify     = new.modify))

set.seed(4)
no_event <- dads(bd.params = bd.params,
                 stop.rule = stop.rule,
                 traits    = traits,
                 modifiers = modifiers)
set.seed(4)
change_spec <- dads(bd.params = bd.params,
                    stop.rule = stop.rule,
                    traits    = traits,
                    modifiers = modifiers,
                    events    = change_speciation)

## Visualise the results
par(mfrow = c(1,2))
plot(no_event, main = "No event")
plot(change_spec, main = "Increase extinction for negative traits\nafter time 3")
```

### Changing branch length when reaching n taxa

```{r}
## Parameters
bd.params <- list(extinction = 0, speciation = 1)
stop.rule <- list(max.taxa = 100)
modifiers <- make.modifiers()
traits <- make.traits()

## multiplying branch length 100 folds
new.modify <- function(x, trait.values, lineage) {
  return(x * 100)
}

## Event for increasing branch length after reaching 30 taxa
increase_brlen <- make.events(
                  condition    = taxa.condition(30),
                  target       = "modifiers",
                  modification = update.modifiers(
                                    branch.length = branch.length,
                                    modify = new.modify))

## Run the simulations
set.seed(5)
no_event <- dads(bd.params = bd.params,
                 stop.rule = stop.rule,
                 traits    = traits,
                 modifiers = modifiers)
set.seed(5)
increased_brlen <- dads(bd.params = bd.params,
                        stop.rule = stop.rule,
                        traits    = traits,
                        modifiers = modifiers,
                        events    = increase_brlen)

## Visualise the results
par(mfrow = c(1,2))
plot(no_event, main = "No event")
plot(increased_brlen, main = "Increase branch length\nafter 30 taxa")
```

### Founding event: a subtree with no fossils

```{r}
## Set up parameters
stop.rule <- list(max.time = 4)
bd.params <- list(speciation = 1, extinction = 0.3)

## Events that generate a new process (founding effects)
founding_event <- make.events(
                  condition    = taxa.condition(10),
                  target       = "founding",
                  modification = founding.event(
                                    bd.params = list(speciation = 2,
                                                     extinction = 0)),
                  additional.args = list(prefix = "founding_"))

## Simulations
set.seed(11)
founding_tree <- dads(bd.params = bd.params,
                      stop.rule = stop.rule,
                      events    = founding_event)
plot(founding_tree, cex = 0.4)
```

### Founding event with traits

```{r}
## Set up parameters
stop.rule <- list(max.time = 4)
bd.params <- list(speciation = 1, extinction = 0.3)
traits    <- make.traits()

## Events that generate a new process (founding effects)
founding_event <- make.events(
                  condition    = taxa.condition(10),
                  target       = "founding",
                  modification = founding.event(
                                    bd.params = list(speciation = 2,
                                                     extinction = 0),
                                    traits = make.traits(process = OU.process)),
                  additional.args = list(prefix = "founding_"))

## Simulating the tree
set.seed(19)
founding_tree <- dads(bd.params = bd.params,
                      stop.rule = stop.rule,
                      traits    = make.traits(),
                      events    = founding_event)
plot(founding_tree)
```