### 2023-01-19 inst/gitbook/05_plotting_dads_objects.Rmd
added ploting tutorial

---
title: "`plot.dads`"
author: "Thomas Guillerme (guillert@tcd.ie)"
date: "`r Sys.Date()`"
bibliography: [../References.bib, ../packages.bib]
output:
  html_document: default
  pdf_document: default
---

```{r, echo = FALSE}
library(rgl)
knitr::knit_hooks$set(webgl = hook_webgl)
```

`"dads"` objects can be directly plotted in dads using many different options.