### 2023-04-05 inst/gitbook/04_make.events.Rmd

---
title: "`make.events`"
author: "Thomas Guillerme (guillert@tcd.ie)"
date: "`r Sys.Date()`"
bibliography: [../references.bib]
output:
  html_document: default
  pdf_document: default
---

```{r, echo = FALSE}
library(rgl)
knitr::knit_hooks$set(webgl = hook_webgl)
```

```{r, echo = FALSE}
library(dads)
set.seed(123)
```

# Adding events to simulations {#makeevents}

One other major feature of the `dads` package is that it allows simulations to run with specified events that occur during the simulation.
These are typically events that can drastically change the course of the simulation.
For example, you might want simulate a mass extinction at some specific point in time.
However, these can also be more subtle, like the internal change of parameter values when reaching a specific trait value.
These are all handled by `dads` with the `events` object that you can create using `make.events` and has the same overall logic as `make.traits` and `make.modifiers`.

`events` require three main arguments:
 * the `target` which designates what the extinction should affect (e.g. the taxa, the speciation rate, etc...);
 * the `condition` which designates when to trigger the event;
 * the `modification` which designates what to modify when the event is triggered.

There are several more arguments that can be passed to `make.events` but they are discussed later on.
First let's focus on these three main arguments:

## Target

The target of the event is what the event is going to modify in the birth-death algorithm.
You can only have one target per event (along with one condition and one modification) but you can create events that contain multiple events (i.e. multiple triplets of target/condition/modification).
The targets that are currently available are:
 * `"taxa"` to modify anything linked to the [`lineage` list](#allowarguments). e.g. making half of the living taxa go extinct.
 * `"bd.params"` to modify anything linked to the [`bd.params` object](#makebdparams). For example you might want to change the distribution of one of the parameter after some `conditions`. This is typically done by updating the object using the argument `update` from `make.bd.params`.
 * `"traits"` to modify anything linked to the [`traits` object](#maketraits). For example you might want to change the trait process after some `conditions`. This is typically done by updating the object using the argument `update` from `make.traits`.
 * `"modifiers"` to modify anything linked to the [`modifiers` object](#makebdmodifiers). For example you might want to change the speciation rule after some `conditions`. This is typically done by updating the object using the argument `update` from `make.modifiers`.
 * `"founding"` this target is a bit more specialised and allows you run a nested `dads` object in the simulation. It is covered in a [specific section below](#founding).

We will see some examples of these targets in the `conditions` and `modifications` described below.

## Conditions

`condition` is a function that returns a logical value. When a specific condition is met, it should return `TRUE` and trigger the event, else it should return `FALSE`.

Currently there are three conditions functions implemented in `dads` but you can easily come up with your own version of them.

All `condition` functions in `dads` take at least two arguments: `x` for the variable of interest (e.g. time, number of taxa, trait value, etc.) and `condition`, the relational operator to evaluate. A relational operator is the proper (fancy) term designating all the comparisons you're regularly using in `R` like `==` (is equal?) `<` (is smaller?) `>=` (is bigger or equal?). You can get the full list in the R base manual (using `?Comparison` or any of the relational operator in a function form). For more details on how functions _really_ work in R see the [Advanced `R` book](https://adv-r.hadley.nz/functions.html)!

 * `time.condition` will trigger the `event` once a certain time is reached in the simulations and is the simplest/most basic condition with no arguments other than the time required. For example ``time.condition(4, condition = `==`)`` will trigger the `event` once the simulations reaches 4 time units. Easy.
 * `taxa.condition` will trigger the `event` once a certain number of taxa are reached. This can be considered including or excluding fossil species. For example ``taxa.condition(42, condition = `>=`, living = TRUE)`` will trigger the `event` if there are at least 42 living taxa.
 * `trait.condition` will trigger the `event` once a certain trait value is reached. This function allows to say which trait(s) to target (by default, the first one using `trait = 1`), what value of the trait to target (by default `what = max`) and whether to use and absolute trait value or not (`absolute = TRUE`). For example ``trait.condition(1/3, condition = `>`, trait = 1, what = sd)`` will trigger the condition after the standard deviation of the first trait reaches 1/3.

## Modifications

After defining the `event`, `target` and `condition`, you also need to specify what it should modify.
These are functions that should modify a specific aspect of the target.
Depending on the target you can modify the following:

 * if the target is `"taxa"` you can modify the internal [`lineage` list](#allowarguments) by removing living species using `random.extinction` or `trait.extinction`.
 * if the target is `"taxa"` you can modify the lineage tracker by removing living species using `random.extinction` or `trait.extinction`.
 * if the target is `"bd.params"` you can modify the [`bd.params` object](#makebdparams) using `update.bd.params`.
 * if the target is `"traits"` you can modify the [`traits` object](#maketraits) using `update.traits`.
 * if the target is `"modifiers"` you can modify the [`modifiers` object](#makemodifiers) using `update.modifiers`.
 * (for the `"founding"` target [see below](#founding))

The most straightforward example is for `modifications` on `bd.params`, `traits` or `modifiers` objects because they use the same syntax as for their generic `make.X` function.
For example, for `make.traits`, you can update a `trait` using the `update` argument as follows:

```{r}
## A BM trait in two dimensions
(BM_2D <- make.traits(n = 2, process = BM.process))
## Updating the 2D BM into a 2D OU
(OU_2D <- make.traits(update = BM_2D, process = OU.process))
```

So the function `update.X` applies the update to the object `X` when the event happens.
The following `event` updates the `bd.params` object by setting the extinction parameter to 1/3 when reaching a 10 species:

```{r, eval = FALSE}
make.events(target = "bd.params",
            condition = taxa.condition(10, condition = `>=`),
            modification = update.bd.params(extinction = 1/3))
```

> Note that by default events are triggered only once across the whole simulation, so although the example above states that the condition is reaching at least 10 taxa, it will not trigger every time it reaches more than 10 taxa, only the first time. You can change the number of times the events can be triggered using the argument `replications` (by default it's set to `replications = 0` for triggering the event only once).

## Examples

Here are some examples illustrating how to generate events.
For a simple example, we can create a extinction event that will remove 80% of species after reaching time 4:

```{r}
## 80% mass extinction at time 4
mass_extinction <- make.events(
                      target = "taxa",
                      condition = time.condition(4),
                      modification = random.extinction(0.8))

## Simulation parameters
stop.rule <- list(max.time = 5)
bd.params <- list(extinction = 0, speciation = 1)

## Running the simulations
set.seed(123)
results <- dads(bd.params = bd.params,
                stop.rule = stop.rule,
                events = mass_extinction)
## Plotting the results
plot(results, show.tip.label = FALSE)
axisPhylo()
```

Or for a slightly more complex example, we can can change the trait process from a BM to an OU when the trait values reaches an upper 95% quantile value above 2:

```{r}
## The 95% upper quantile value of a distribution
upper.95 <- function(x) {
    return(quantile(x, prob = 0.95))
} 
## Create an event to change the trait process
change_process <- make.events(
                  target = "traits",
                  ## condition is triggered if(upper.95(x) > 3)
                  condition = trait.condition(3, condition = `>`, what = upper.95),
                  modification = update.traits(process = OU.process))

## Set the simulation parameters
bd.params <- list(extinction = 0, speciation = 1)
stop.rule <- list(max.time = 6)
traits    <- make.traits()

## Run the simulations
set.seed(1)
no_change <- dads(bd.params = bd.params,
                  stop.rule = stop.rule,
                  traits = traits)
set.seed(1)
process_change <- dads(bd.params = bd.params,
                       stop.rule = stop.rule,
                       traits = traits,
                       events = change_process)
## Plot the results
par(mfrow = c(1,2))
plot(no_change, ylim = c(-7, 7))
plot(process_change, ylim = c(-7, 7))
par(mfrow = c(1,1))
```

## Founding events {#founding}

Founding events are a specific `target` for `events` that allows you to simulate a birth-death process within the current one!
It allows you to simulate a specific `dads` process using the `dads` function with its own `traits`, `modifiers`, `bd.params` and `events` (and yes, that's including `events` that have their own `founding` event).
This is basically the ultimate nested boss of modularity!

The founding event will run an internal `dads` process (i.e. simulating a tree and, optionally, some data) resulting in a founding sub-tree.
It will then branch this sub-tree onto the rest of the simulation that continued normally in the mean time.
You can specify the founding event using the inbuilt `founding.event` modification event.
This function takes the exact same arguments as `dads` to simulate the sub-tree with its own parameters.
Additionally, we will use the `additional.args` argument from `make.events` to specify a prefix for the founding tree tips (to make them easier to distinguish).

```{r}
## Set up parameters
stop.rule <- list(max.time = 4)
bd.params <- make.bd.params(speciation = 1, extinction = 0.3)

## Events that generate a new process (founding tree - with no extinction)
founding_event <- make.events(
                  target = "founding",
                  condition = taxa.condition(10),
                  modification = founding.event(
                                    bd.params = make.bd.params(speciation = 2,
                                                               extinction = 0)),
                  additional.args = list(prefix = "founding_"))

## Simulations
set.seed(11)
founding_tree <- dads(bd.params = bd.params,
                      stop.rule = stop.rule,
                      events = founding_event)
plot(founding_tree, cex = 0.4)
```

Note that the nestedness here is potentially endless, for example, you can pass an `event` argument to the `founding.event` that will generate another founding event, etc.