### 2023-04-24 inst/gitbook/verbose_example.Rmd

---
title: "`treats` verbose example"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

```{r}
library(dispRity)
library(treats)
```

```{r}
## K-Pg mass extinction in mammals
data(BeckLee_tree)
data(BeckLee_mat99)
data(BeckLee_ages)
time_series <- chrono.subsets(data = BeckLee_mat99, tree = BeckLee_tree,
                              method = "continuous", model = "acctran", time = 20,
                              FADLAD = BeckLee_ages)
result <- dispRity(time_series, metric = c(sum, variances))
plot(result)
abline(v = 10)
results <- dispRity(boot.matrix(time_series), metric = c(sum, variances))
plot(results)
abline(v = 10)
```

```{r}
libary(treats)
my_params <- make.bd.params(speciation = rnorm)
my_stop_rule <- list(max.taxa = 50)
trait <- make.traits(process = BM.process, n = 10)
simulation <- treats(bd.params = my_params, stop.rule = my_stop_rule, trait = trait)

## Measure disparity through time
time_series <- chrono.subsets(data = simulation$data, tree = simulation$tree,
                              method = "continuous", model = "acctran", time = 20)
results <- dispRity(boot.matrix(time_series), metric = c(sum, variances))
plot(results)
```

```{r}
## 
library(treats)
set.seed(123)
bd.params <- make.bd.params(speciation = 1)
stop.rule <- list(max.taxa = 50, max.living = Inf, max.time = Inf)
traits <- make.traits(process = BM.process, background = make.traits())
par(mfrow = c(2,1))
plot(treats(bd.params = bd.params, stop.rule = stop.rule, traits = traits), main = "no modif")

## We want to change this in two ways:
 # 1- speciation = trait.value^2

select.scale.to.absolute.trait.value <- function(trait.values, lineage) {
    ## Sampling any of the living lineages with the probability of being selected being relative to the current lineage trait value. The bigger the value, the higher the probability

    ## Get the trait values of the descendants (as base probabilities + 0.01 to avoid 0 probabilities)
    sample_probs <- (abs(trait.values[as.character(lineage$parents[lineage$livings]), ]) + 0.01)

    return(as.integer(sample(1:lineage$n, size = 1, prob = sample_probs )))
}

modifiers <- make.modifiers(selection = select.scale.to.absolute.trait.value)

plot(treats(bd.params = bd.params, stop.rule = stop.rule, traits = traits, modifiers = modifiers), main = "Speciation increase\nwith absolute trait value")

## THIS IS BUGGED: to test in birth.death_fun.R: problem with selecting the current lineage (NA)

```