### 2023-05-26 paper/treats_paper_plots.Rmd
Create treats_paper_plots.Rmd

---
title: "treats paper plots"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
bibliography: references.bib
output:
  html_document:
    fig_width: 8
    fig_height: 8
---

This script reproduces the plots and analyses used in the `treats` paper.

## Beck Lee data (figure 1)

```{r}
## Loading the data and packages
library(treats)
library(dispRity)
data(BeckLee_tree)
data(BeckLee_mat99)

## Creating the time slices
time_slices <- chrono.subsets(BeckLee_mat99, BeckLee_tree,
                              method = "continuous",
                              model  = "acctran",
                              time   = seq(from = 96, to = 36, by = -10))

## Calculating disparity on the two first dimensions only
observed_disparity <- dispRity(boot.matrix(time_slices), metric = c(sum, variances), dimensions = c(1,2))

## Plotting the tree and the disparity through time
pdf(file = "treats_example_figure1.pdf", width = 14, height = 7)
par(mfrow = c(1,2))
plot(BeckLee_tree, show.tip.label = FALSE)
axisPhylo()
abline(v = BeckLee_tree$root.time - 66, col = "red")
legend("topleft", pch = NULL, legend = "A", bty = "n", cex = 2)
par(bty = "n")
plot(observed_disparity, ylab = "Sum of variances")
abline(v = 4, col = "red")
legend("topleft", pch = NULL, legend = "B", bty = "n", cex = 2)
dev.off()
```