### 2023-08-11 paper/response.Rmd
response INIT

---
title: "Response to reviewers"
author: "Thomas Guillerme"
bibliography: References.bib
# csl: mee.csl
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    keep_tex: yes
    fig_width: 8
    fig_height: 8
  bookdown::html_document2: default
  bookdown::pdf_document2: default
---

Dear Editor,

Please find my response to the reviewers comments in the following document.
I've highlight <span style="color:blue">the Associate Editor and both reviewers comments in blue</span> and my response black.

Best regards,
Thomas

# Associate Editor Comments to Author:

<span style="color:blue">As both reviewers, I really liked the idea of the package, but also found the manuscript itself a bit hard to follow, which, at least to me, made it hard to fully appreciate what “treats” can currently do and what are the future additions that might be easily implemented (either by the developer, or by users). Reading the online manual, I got a better idea, and hence I think there is still room (metaphorically and I suspect also physically) for improvement. Sure, the online material has no space limit constraints, but I do think the manuscript text could be improved and both reviewers offered some nice suggestions.  Apart from the really good suggestions given by both reviewers I would add the following suggestions/comments:</span>

* <span style="color:blue">1- Are the trait values “stored” only at the node and tip (extant or extinct), or does it store (e.g. assuming some small-time intervals) the trait values throughout the branch length of each lineage? Based on lines 63 to 65 (and 71) I think only at nodes and tips, but it might be worth explicitly saying that. Perhaps add the word “only” in that line.  Also, in line 71 is there an extra “no” within the parenthesis?</span>

* <span style="color:blue">2- Regarding the “storage” of trait value, if it only stores the values at nodes and tips, does it also mean that it only uses trait values at those time points to direct the simulation? Does that matter? What made me wonder about that was figure 5, and its respective intended selective extinction simulation scenario. If I got it right, the selective extinction scenario would remove all species with positive trait values at the mass extinction event at 15 Mya. Looking at the right panel, all species whose direct ancestral node had a positive trait value seemed to be removed, but some of those with ancestral negative values likely crossed the red line having a positive value (even if the “true trait trajectory” is not a straight as represented, for simplicity I guess, in the figure) and were not removed. This might not be always relevant, and I understand that some simplifications are necessary, but it might be worth explicitly discussing if this is relevant and when. For example, if speciation and background extinction rates are small, species are likely to have long branches and the trait value it might have at a given time point between its ancestral node and its tip value or next speciation node might be very different from the value it had at the ancestral node (or tip value for that matter). Or is the decision to remove the species in the selective mass extinction event one where both the ancestral and tip value have to be positive. I guess some extra information on how it chooses the species based on the trait value (is it the ancestral node value, the tip value, both?) in the selective mass extinction regime would also be helpful. I guess the issue seems that the trait value is not measured at short time intervals (only at the nodes and tips) but a selective extinction regime such as this one presented here is a point in time event. </span>

* <span style="color:blue">3- I have to say that looking at figure 5 it was not trivial to see the difference between the two scenarios (I guess that happened because most, if not all, species with positive values went extinct in the left random panel as well). I know those are realizations of a stochastic process and there is nothing wrong here, my concern regards simply the visualization. I wonder if choosing another seed might result in a set of simulations that are more clearly different, making the comparison easier for the eye to pick it up. For example, in the online manual, some examples comparing random to selective simulations showed a much more disparate pattern.</span>

* <span style="color:blue">4- Code box on page 7: the commentary on the second simulation scenario reads “#Simulate the tree and traits with a random extinction event” and hence is equal to the comment in the first simulation scenario in the same box. Perhaps change it to something like: “#Simulate the tree and traits with a random extinction event but based on trait value”</span>

* <span style="color:blue">5- I could be wrong but it seems that figure 6 is not cited in the text. Please double check and if this is true, please cite this figure. </span>

* <span style="color:blue">6- All figure captions have some repetition like “Figure 1: Figure 1:….”</span>

* <span style="color:blue">7- Figure 6: I could be wrong, but my intuition was that a selective mass extinction event would decrease the disparity more strongly than a random extinction mass extinction event. That does not seem to be the case. Is it because the “rule” for the second extinction scenario was that only positive species will go extinct, hence on average 50% of the species (if you start with a value of zero) instead of 75% went extinct? That would make sense, I guess. Or is the second extinction scenario one where you also create a 75% loss of species but guarantee that all species with positive values will go extinct? Maybe a few words to better explain this would be helpful. Perhaps add on lines 109 to 110 something like: “note that in the second scenario, because the ancestral state is zero and we do not set the percentage of species to be removed, we expect only 50% of the species to go extinct”. Just an idea.</span>

# Reviewer: 1

<span style="color:blue">This article by Guillerme introduces a new R package treats, which implements a flexible framework for simulating phylogenies and associated trait values. Overall, I really like the concept of the package, and I think it can be very useful for more advanced users who run many simulations and change their setup often. My comments are mostly about clarifying the presentation of the different functions. I also reviewed the manual.</span>

## Comments on the manuscript

* <span style="color:blue">I had trouble understanding what treats could and could not do from the manuscript alone. The introduction discusses several packages related to fossil data (FossilSim, paleotree and paleobuddy) which gave me the impression that treats could replace them, but it doesn't seem that fossilization processes are currently included ? I was also a bit confused by the inclusion of time-dependent and lineage-dependent simulations in future features, as it seems that the current framework should be able to handle these, at least within the birth-death process. Overall, having a clearer list of what is currently possible or not in treats would be helpful.</span>

* <span style="color:blue">I would also have liked some information on the performance of the package compared to existing simulation tools. It seems likely that the flexibility offered by treats comes at the cost of some computational performance, so it would be nice to know how much time would be needed to simulate a reasonable dataset (for instance the 50 replicates used to test the positive extinction vs random extinction).</span>

* <span style="color:blue">I didn't understand why the example shown in the manuscript only simulated trees up to 30 time units, when the tree used as a basis for the comparison is 140 My old (figure 2).</span>

* <span style="color:blue">I really like the idea of having a library of premade simulation templates (l202) and I believe it will be a huge help in making the package accessible. Making the examples shown in the manual (section 8) available as scripts would be a good start for such a library.</span>

### Minor comments:

* <span style="color:blue"> - l71 "no either" </span>

* <span style="color:blue"> - l125 "where the it"</span>

* <span style="color:blue"> - l111 the comment in the second example is mismatched with the code (random extinction instead of positive extinction)</span>

* <span style="color:blue"> - l175 the comment in the example is mismatched with the code (positive extinction instead of random extinction)</span>

* <span style="color:blue"> - l178 the text mentions the function trait.extinction but it's not used in the code</span>

* <span style="color:blue"> - l282 missing authors in the reference</span>

## Comments on the manual

* <span style="color:blue">The main thing the manual is missing in my opinion is a quick reference table of the different functions, for people who have understood the concepts of the package but need to quickly check what format their custom speciation function needs to follow for instance. At the moment that information is incomplete (the table in section 4.5 shows the output types but not what these outputs are) and spread out in different places (the description of the function inputs is in section 4.2.1 whereas the list of functions is in 4.1).</span>

* <span style="color:blue">The use of conditions and modifications was quite confusing to me. Some of the examples use the modify and condition argument to change the birth-death process (section 2.3) whereas some examples simply replace the function with a modified version (section 4.4), and I could not figure out if these methods were equivalent or if one should be preferred to the other. I think the modifications section also needs more details about what exactly is being modified, and what the expected inputs and outputs of the modification function should be. For instance, in the example section 2.3 I don't understand what is the "x" argument in "going.extinct", or why this function returns a number instead of a logical. The interface for the conditions in 5.2 appears to be clearer, but it doesn't match with the example in section 2.3 where the "negative.ancestor" uses different arguments, or with the example in 4.6 where the "half.the.time" condition takes no arguments. </span>

* <span style="color:blue">Finally, I also could not find a full description of the treats output object either in the manual or in the package documentation.</span>

# Reviewer: 2

<span style="color:blue">Thank you for the opportunity to review this manuscript, titled “treats: a modular R package for simulating trees and traits”. This manuscript describes and showcases a new R package that provides users with an extensive toolbox to jointly simulate phylogenies and evolving traits. While the example provided within the manuscript gives readers a small taste of the flexibility of this toolbox, the online manual (http://tguillerme.github.io/treats.html) showcases the true extent of how useful this package will be for evolutionary biologists far and wide. I really enjoyed reading the manuscript, and I believe it will be suitable for publication in Methods in Ecology & Evolution once my concerns have been addressed.</span>

<span style="color:blue">For context, for this review I read the supplied manuscript, tested the functionality of the R package, and browsed the built-in documentation and the online supplemental manual. I was unable to review the functional code of the R package in time for the due date of this review, but it appears to work properly from my testing. I have one major concern and a handful of minor concerns. The major concern is that the overall flow of the manuscript seems a little off. I think this mostly stems from the worked example preceding the detailed overview of the package’s functions. I think if these two sections were swapped the manuscript would flow much better. My minor concerns are included at the end of this review in order of their appearance in the manuscript. I look forward to reviewing a revised version of this manuscript, if necessary.</span>

<span style="color:blue">Best,</span>

<span style="color:blue">William Gearty</span>

* <span style="color:blue">Lines 27-31: I would enumerate this list.</span>

* <span style="color:blue">Line 84: Geological time period names should be capitalized.</span>

* <span style="color:blue">Figures 2-6: If possible, I think it would be valuable to show what code was used to produce these figures in the code snippets. As far as I can tell, the plotting is mostly done using functions from the package, so this would also be an opportunity to showcase this functionality.</span>

* <span style="color:blue">Figure 2: The caption should identify what the red line represents (the K-Pg boundary?).</span>

* <span style="color:blue">Line 111: This may be personal preference, but when I ran this code chunk it output a whole bunch of text to the console. It appears to be a separate line of text for each replicate in the treats function. I believe the intent here is to show the user how much progress has been made? Perhaps some sort of progress bar would be more appropriate?</span>

* <span style="color:blue">Lines 106-111: I think you should mention somewhere in this paragraph why you are running multiple replicates here (as opposed to the single replicate performed earlier in the worked example).</span>

* <span style="color:blue">Line 111: I used the same seed as you, but I was unable to reproduce Figure 5 (looking at the R markdown file, I see that you use indices 11 and 10 to make the plots). I think showing the exact code that is used here to make Figure 5 would be useful for readers that might be “playing along at home”.</span>

* <span style="color:blue">Lines 112-113: You should refer to Figure 6 somewhere in this paragraph. It might also be useful to include the code to perform this comparison, especially if the functionality is all contained within the treats package.</span>

* <span style="color:blue">Line 194: This reference to the manual should have a link to the online manual (looks like the link is embedded, but you should also have it typed out). I also noticed that the documentation for the treats function lists the wrong URL for the online manual.</span>

* <span style="color:blue">Lines 202-203: Just wanted to note that I really love this idea! I think collecting a library of modules will be very useful to users!</span>

* <span style="color:blue">Line 205: This embedded link does not work. I think it should be something like: https://github.com/TGuillerme/treats/tree/master/paper.</span>