### 2024-01-15 inst/ace_treats.Rmd
basic (not good!) ace

---
title: "Running ancestral states estimations on treats"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 6
    fig_height: 6
---

## Ace pipeline idea

1 - estimate the root (x0)
2 - estimate the parameters (e.g. variance)
3 - simulate the process
4 - test the fit with the observed data.

## Readings

https://link.springer.com/article/10.1007/s11692-020-09505-x

Cunningham, C. W., Omland, K. E. and Oakley, T. H. (1998) Reconstructing ancestral character
states: a critical reappraisal. Trends in Ecology & Evolution, 13, 361–366

Pagel, M. (1994) Detecting correlated evolution on phylogenies: a general method for the com-
parative analysis of discrete characters. Proceedings of the Royal Society of London. Series B.
Biological Sciences, 255, 37–45

## Playground

```{r}
library(treats)
set.seed(1)
play_data <- treats(bd.params  = list(speciation = 1, extinction = 0.5),
                    traits     = make.traits(),
                    stop.rule  = list(max.taxa = 50),
                    null.error = 100) 
plot(play_data)
```

https://github.com/pasraia/RRphylo/blob/master/R/RRphylo.R

https://github.com/emmanuelparadis/ape/blob/master/R/ace.R